\documentclass[a4paper,11pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{array}
\usepackage{longtable}
\usepackage{fancyhdr}
\usepackage{amssymb}  % for checkmarks and crosses
\usepackage{xcolor}   % for coloring the symbols
\usepackage{pifont}   % for coloring the symbols
\usepackage{enumitem}
\setlist[itemize]{topsep=0pt, partopsep=0pt, itemsep=0pt, parsep=0pt}

\newcommand{\done}{\textcolor{green}{\ding{52}}}
\newcommand{\ongoing}{\textcolor{orange}{\ding{45}}}
\newcommand{\notstarted}{\textcolor{red}{\ding{56}}}


\raggedright

\pagestyle{fancy}
\fancyhf{}
\lhead{Bachelor Thesis Update}
\chead{Nadja Stadelmann}
\rhead{\today}
\renewcommand{\headrulewidth}{0.4pt}

\begin{document}
	
	\renewcommand{\arraystretch}{1.5}
	\begin{longtable}{|p{3.5cm}|>{\vspace{-\baselineskip}}p{12cm}|}
		\hline
		\textbf{Section} & \textbf{Details} \\
		\hline
		Progress Overview (inkl. Reasons for Deviations of the Plan) & \begin{itemize}[topsep=6pt]
			\item[\done] Evaluated and fixed PASSION results reproduction issue on group level
			\item[\done] Fixed checkpoint handling
			\item[\ongoing] Code integration into PASSION
			\begin{itemize}
				\item I would need to test (and potentially fix) it also for the impetigo experiment for a proper integration
			\end{itemize}
			\item[\done] Ran the evaluation on the tiny model to get a proper baseline
			\item[\done] Analyzed which biases exist in the baseline
			\item[\ongoing] Stratified split experiment
			 \begin{itemize}
			 	\item[\done] Created split files using different stratified splits based on identified biases
			 	\item[\done] Trained small models with all split files without folds
			 	\item[\notstarted] Analyze runs, rerun with folds for the 2 best performing split files to evaluate the best split
			 \end{itemize}
			\item[\done] Progress on the report
			\item[\done] Created a plan for the last 3 weeks
			\item[\notstarted] another experiment, due to missing time
		\end{itemize} \\
		\hline
		Accomplishments & \begin{itemize}[topsep=6pt]
		\item[\done] Provided value by fixing the code
		\item[\done] Bias analysis on tiny model baseline
		\item[\done] First results regarding stratified split experiment ready for analysis
		\end{itemize} \\
		\hline
		Risks and Measures &
		
		Incomplete code integration leading to loose strings at the end of the thesis
		\begin{itemize}
			\item Focus on finalizing the integration instead of starting new experiments
			\item OR: officially delay the proper code integration until after the thesis
		\end{itemize}
		
		Burn out
		\begin{itemize}
			\item Reduce external workload by taking time off from work (ongoing).
			\item Focus on finalizing existing work.
			\item Schedule and take explicit off time.
		\end{itemize}
		
	    Deliver too less in terms of executed experiments
		\begin{itemize}
			\item Transparently explain the trade-offs made in regards of the quality-over-quantity approach.
			\item Outline which further experiments could be done in the outlook.
			\item OR: do another quick experiment
		\end{itemize} \\
		
		\hline
		Next Steps & \begin{itemize}[topsep=6pt]
			\item See attachement
		\end{itemize} \\
		\hline
		Discussion Points / Questions & \begin{itemize}[topsep=6pt]
			\item What are the expectations for the pitch video besides that it takes 1.5 minute?
			\item Do you expect another report for the meeting in the last week?
		\end{itemize} \\
		\hline
		Additional Notes & - \\
		\hline
		Attachments  & \begin{itemize}[topsep=6pt]
			\item Plan for the upcoming weeks
		\end{itemize} \\
		\hline
		Protocol  & 
		Topic
		\begin{itemize}
			\item 
		\end{itemize}
		
		Topic
		\begin{itemize}
			\item 
		\end{itemize}
		\\
		\hline
	\end{longtable}
	
\end{document}
